---
title: "manuscript"
format: html
---

## Background
This analysis is based on the paper by Shaefer et. al. 2000 "Effects of hurricane disturbance on stream water concentrations and fluxes in eight tropical forest watersheds of the Luquillo Experimental Forest, Puerto Rico". In this study they assessed effects of tropical hurricanes on stream water chemistry in various watersheds of the Luquillo Experimental Forest (LEF) in Puerto Rico. They specifically looked at the effects after the 1989 Hurricane Hugo.

In this analysis, we are creating a workflow to recreate Figure 3. This plot shows the ion concentrations for 5 ions (NH4-N, Ca, Mg, NO3-N, and K) over time, separated by the four Bisley, Puerto Rico streams. 

## Data
These data and associated metadata can be found here https://doi.org/10.6073/PASTA/F31349BEBDC304F758718F4798D25458. These data run from 1986-2020, with weekly water sampling for various ion concentrations. This analysis consists of only four stream sites: MPR, Q1, Q2, and Q3. These are all from the Bisley, Puerto Rico watershed area. Chemical concentrations are recorded as mg/L or mg/L as appropriate. Values below detection limits are recorded as 1/2 the detection limit.

## Methods
These concentrations are calculated as 9 week moving averages.

## Results
```{r}
#| echo: false
# read in packages
library(here)
library(tidyverse)
# source the plot function
source(here("R", "plot.R"))
# read in clean data
full_data <- read_csv(here("output", "full_data.csv"))
```

```{r}
# apply the plot function with the data and parameters
hurricane_ions_plot(full_data, sample_date, concentration)
```

```{r}
#| echo: false
# saves the final plot into the figures folder
ggsave(here("figs", "figure-3.jpg"))
```

